---
title: "TidyTuesday 12/09/2024"
author: "Megan Smith"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

Load Libraries
```{r}
library(here)
library(tidyverse)
library(tidytuesdayR)
library(plotly)
library(RColorBrewer)
```

Bring in data 
```{r}
tuesdata <- tidytuesdayR::tt_load('2024-11-26') # pulling in data from tidy tuesday
```
Subsetting the data 

```{r}
cbp_resp <- tuesdata$cbp_resp
cbp_state <- tuesdata$cbp_state
```

Glimpsing the data 
```{r}
glimpse(cbp_resp)
glimpse(cbp_state)
```
Gor this data set I want to look at total encounters by year and by demographic. 

Cleaning data 
```{r}
combined_data <- bind_rows(cbp_state, cbp_resp) # combining the data sets 

demographic_data <- combined_data %>% # making a demographic data set
  group_by(fiscal_year, demographic) %>% # grouping year and demographic
  summarise(total_encounters = sum(encounter_count, na.rm = TRUE)) %>%  # summarizing total encounters by the grouping 
  mutate(demographic = recode(demographic, "FMUA" = "Family Unit")) # making this demographic variable a little more readable
```



# Something new

This is my first time using the package "Plotly" where you can hover over th figure and it is interactive. 

 For my purposes I used it so you could hover and see the exact number of encounters for each facet and for each year. 

```{r encounters, dev"png"}
p <- ggplot(demographic_data, aes(x = as.factor(fiscal_year), # making year a factor so it goes in order
                                  y = total_encounters, 
                                  fill = demographic, 
                                  text = paste("Total Encounters:", scales::comma(total_encounters))
                                  )) + # added this so when you hover it only shows total and with commas encounters
  geom_bar(stat = "identity", position = "dodge") + # how bar plot will how
  facet_wrap(~ demographic, scales = "free_y") + # faceting by demographic with free y axis
  labs(title = "Total Encounters by Demographic and Fiscal Year", # creating new label names 
       x = "Fiscal Year", 
       y = "Total Encounters",
       fill= "Demographic") +
  scale_fill_brewer(palette = "Dark2")+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #angeled x axis labels 


interactive_plot <- ggplotly(p, tooltip = "text") # tells that it only wants text appeared from above so it only shows total encounters and with the commas
interactive_plot # to dispay the figure
```

