---
title: 'Tidy Tuesday: Week 45'
author: "Megan Smith"
date: "2024-11-05"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

# Load Libraries 
```{r}
library(tidyverse)
library(here)
library(maps)
```

# Load in Data 
```{r echo = T, results = 'hide'}
tt_1105 <- tidytuesdayR::tt_load('2024-11-05')

dem_data<-tt_1105$democracy_data

glimpse(dem_data)
```
# Cleaning Data

 - I want to see the average age of accesion by country

```{r}
dem_data <- dem_data %>%
  mutate(pres_age_at_acc = if_else(!is.na(president_accesion_year) & !is.na(president_birthyear), # creating new column for age at acc
                                         president_accesion_year - president_birthyear, #accesion year - birth year to come up with age at succession 
                                         NA_real_),
         mon_age_at_acc = if_else(!is.na(monarch_accession_year) & !is.na(monarch_birthyear),# same for monarchs
                                       monarch_accession_year - monarch_birthyear,
                                       NA_real_))
```

- Now getting the mean age of each country
```{r}
avg_age_acc <- dem_data %>%
  group_by(country_name) %>% # grouping to get average for each country
  summarize(avg_pres_age_at_acc = mean(pres_age_at_acc, na.rm = TRUE),
            avg_mon_age_at_acc = mean(mon_age_at_acc, na.rm = TRUE))
```


# Calling in Map Data 

 
```{r}
world <- map_data("world")
```

- Combining president and monarch average age of accession 
```{r}
avg_age_acc <- avg_age_acc %>%
  mutate(avg_age_of_accession = rowMeans(select(., avg_pres_age_at_acc, avg_mon_age_at_acc), na.rm = TRUE)) # rowMeans to calc means across columns 
```

- Changing country_name to region to match world data
```{r}
avg_age_acc <- avg_age_acc %>%
  select(region =country_name, avg_age_of_accession)
```

- Attempting to match some of the country names to match world data 
```{r}
avg_age_acc <- avg_age_acc %>%
  mutate(region = recode(str_trim(region), # changing names 
                         "United States" = "USA",
                         "United Kingdom" = "UK",
                         "Korea (Rep.)" = "South Korea",
                         "Congo (Dem. Rep.)" = "Democratic Republic of the Congo",
                         "Congo (Rep.)" = "Republic of Congo",
                         "Saint Barthélemy" = "Saint Barthelemy",
                         "Côte d'Ivoire" = "Ivory Coast")) 
```

# Merging World Data with Dem Data
```{r}
mapmerge<-inner_join(world,avg_age_acc, by= "region")

```

# Map Figure

- Something new!
  - I have not made a plot like this using the world data 
   - [Here](https://sarahpenir.github.io/r/making-maps/) is where I learned how to make the below figure
```{r avg_age_accession_map, dev= "png",fig.width=8, fig.height=6}
plain <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(hjust = 0.5)
)
ggplot(data = mapmerge, mapping = aes(x = long, y = lat, group = group)) +  # cant forget group=group
  coord_fixed(1.3) +
  geom_polygon(aes(fill = avg_age_of_accession)) + # fill with average age 
  scale_fill_distiller(palette ="RdBu", direction = -1) + # from above site 
  labs(title= "Average Age of Accession by Country", # changing title 
      fill="Average Age",
      caption= "Grey=NA")+ #changing legend title
  plain # adding above code
```

