---
title: "TinyTuesday12/3"
author: "Megan Smith"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

Load Libraries
```{r}
library(tidytuesdayR)
library(tidyverse)
library(here)
library(ggmap) #for ggmaps
library(hrbrthemes)
library(paletteer)

```

Pulling in Tidy Tuesday Data 
```{r}
tuesdata <- tidytuesdayR::tt_load('2024-12-03')

A64_traffic <- tuesdata$A64_traffic
```

Exploring Data
```{r, results= "hide"}
glimpse(A64_traffic)
unique(A64_traffic$SiteId)
```


Making map of location of Road Sensors
```{r road_sensor_map,dev="png"}
site<-data.frame(lon= -0.6946696, lat= 54.1581) # coordinates of location
londmap<-get_map(site, zoom= 10) # creating map with those coordinates
ggmap(londmap)+ #map layer
  geom_point(data=A64_traffic, 
             aes(x=Longitude, y=Latitude), # using geom point to mark road sensors
             size=3)+ #size of point
  labs(x="Longitude", #relabeling labels 
       y="Latitude",
       title="Location of Road Sensors")
```




Cleaning Data to Create the Figure
```{r}
A64_traffic <- A64_traffic %>% select(-"Avg mph") # remove 'Avg mph' column

speed_cols <- names(A64_traffic)[grep("mph", names(A64_traffic))] # identify columns with 'mph'

speed_data <- A64_traffic %>%
  pivot_longer(cols = all_of(speed_cols), # pivot these columns into long format
               names_to = "Speed", 
               values_to = "Amount") %>%
  filter(!is.na(Amount)) # remove rows with NA in the 'Amount' column

```

Creating Speed Distribution Plot

 For my something new, I used a new to me color pallete. 
  [On this](https://r-graph-gallery.com/color-palette-finder) site it allows you to filter out color palettes  based on multiple different search criteria. For this figure I needed a lot of colors, so I filtered out a palette that had 14 colors. 
```{r speed_data, dev="png"}
ggplot(speed_data, aes(x = `Report Date`, y = Amount, fill = Speed)) + # data I want to plot 
  geom_bar(stat = "identity") + # selecting bar chart
  scale_fill_paletteer_d("impressionist.colors::korenveld_onder_onweerslucht") + # new color palette from above 
  theme_bw() + #bw theme 
  labs(title = "Speed Distribution Over Time", # changing the labels 
       x = "Date", 
       y = "Total Cars Passing Sensor") 


```



