---
title: "Week 8b Lecture"
author: "Megan Smith"
date: '2024-10-18'
output: rmdformats::downcute
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

# Week 8b Lecture: Advanced plotting 

```{r}
library(tidyverse)
library(here)
library(patchwork)
library(ggrepel)
library(gganimate)
library(magick)
library(palmerpenguins)
```

# Patchwork: easily bring plots together
 these would be plots that cant be mixed using facet
```{r}
# plot 1
p1<-penguins %>%
  ggplot(aes(x = body_mass_g, 
             y = bill_length_mm, 
             color = species))+
  geom_point()
p1
```
```{r}
p2<-penguins %>%
  ggplot(aes(x = sex, 
             y = body_mass_g, 
             color = species))+
  geom_jitter(width = 0.2)
p2
```
```{r}
p1 +p2
```

```{r}
p1+p2 +
  plot_layout(guides = 'collect') #this just makes one lengend (guides same as legend)
```
Adding annotations **this will be helpful**
```{r}
p1+p2 +
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')
```
Stacked instead of next to each other
```{r}
p1/p2 +
  plot_layout(guides = 'collect')+
  plot_annotation(tag_levels = 'A')
```
https://raw.githack.com/OCN-682-UH/Fall_2024/main/Week_08/16_AdvancedPlotting.html#11

#ggrepel: easy and clear labels for plots
Using mtcars thats already part of tidyverse

mtcar plot: labels are on top of each other
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) + #taking row names for points 
  geom_text() + # creates a text label
  geom_point(color = 'red')
```
geom_text_repel
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_text_repel() + # repel them
  geom_point(color = 'red')
```
geom label puts a box around
```{r}
ggplot(mtcars, aes(x = wt, 
                   y = mpg, 
                   label = rownames(mtcars))) +
  geom_label_repel() + # repel them
  geom_point(color = 'red')
```
More on grepple: https://ggrepel.slowkow.com/articles/examples.html

# gganimate: animates figures
More examples: https://gganimate.com/index.html
This is our basic plot with all the years
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()
```

Have to add transition so it knows what to move 
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )
```
Adding more aes to plot: Below is bounce in out 
Other options are linear (default), sine, circular
```{r}

penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  ease_aes("bounce-in-out")
```

Adding labels to transition
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  ease_aes("sine-in-out") +
  labs(title = 'Year: {closest_state}')
```
 Saving as agif -> anim_save
 
 More gganimate here:https://raw.githack.com/OCN-682-UH/Fall_2024/main/Week_08/16_AdvancedPlotting.html#24
```{r}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point() +
  transition_states(
     year, # what are we animating by
     transition_length = 2, #The relative length of the transition.
     state_length = 1 # The length of the pause between transitions
   )+
  ease_aes("sine-in-out") +
  labs(title = 'Year: {closest_state}') +
  anim_save(here("Week_08","output","mypengiungif.gif"))
```

#magick: advanced image processing **this will be very helpful for my prohect**
image_read= is how you will bring the image in 
```{r}
penguin<-image_read("https://pngimg.com/uploads/penguin/pinguin_PNG9.png")
penguin
```
Now to put it on plot- have to save as png
```{r,penguinplot,dev="png"}
penguins %>%
ggplot(aes(x = body_mass_g, 
            y = bill_depth_mm, 
            color = species)) +
  geom_point()
```
Whatever you want as the background goes first
image_composite is how you put them on top: backgroud goes first, second layer after 

```{r}
penplot<-image_read(here("Week_08","output","penguinplot.png"))
out <- image_composite(penplot, penguin, offset = "+70+30") #offset is where you want the penguin to be
out
```

```{r}
pengif<-image_read("https://media3.giphy.com/media/H4uE6w9G1uK4M/giphy.gif")
outgif <- image_composite(penplot, pengif, gravity = "center")
animation <- image_animate(outgif, fps = 10, optimize = TRUE)
animation
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}


```

