---
title: 'Tidy Tuesday: Week 44'
author: "Megan Smith"
date: "2024-11-02"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

# Monster Movie

 This data goes from the [Monster Movies](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-10-29/readme.md) data set. 
 
# Load libraries
```{r}
library(tidyverse)
library(here)
library(naniar) # new package that shows missing data 
```
 
# Loading data
```{r}
tt_10_29 <- tidytuesdayR::tt_load('2024-10-29')
tt_genres<- tt_10_29$monster_movie_genres
tt_movies<-tt_10_29$monster_movies
```
 
# Exploring data 

```{r}
glimpse(tt_movies) #just glimpsing this data since I wont use the other one
```
# New package 

  [Naniar](https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html) is a package that allows you to visualize missing data 
  
  Now I know where missing data is and I can continue cleaning.
```{r visual_missing, dev= "png"}
gg_miss_var(tt_movies) #super cool
```

# Cleaning Data

 I have decided to look at genre type from 2000-2024.
 
  - First I will take only the first genre listed as main genre
  - Then I will create a "True Horror" list to just keep the spooky stuff
  - Merge similar genre types
  
```{r}
movies_clean <- tt_movies %>%
  filter(!is.na(genres)) %>% #removing NAs
  mutate(main_genre = sub(",.*", "", genres)) #genres have more than one, this is selecting just the first as the "main" genre

true_horror<-c("Horror", "Crime", "Documentary","Thriller","Sci-fi","Mystery","Drama","Biography","Adventure","Action") #selecting only what I believe is true horror

genre_year<-movies_clean %>% 
  filter(year >1999) %>% # filtering out 2000 and above
  select(year,main_genre) %>% # selecting what I need
  filter(main_genre %in% true_horror) %>% # filtering our non horror
  mutate(main_genre = case_when( # merging similar genres
    main_genre %in% c("Action","Adventure") ~ "Action-Adventure",
    main_genre %in% c("Crime", "Drama") ~ "Crime-Drama",
    main_genre %in% c("Biography", "Documentary") ~"Biography-Documentary",
    main_genre %in% c("Mystery","Thriller") ~"Mystery-Thriller",
    main_genre == "Horror" ~ "Horror")) %>% 
  group_by(year) # grouping by year
```

# Prepairing for plotting 

 I want to make a proportion stacked bar plot, so I need to wrangle the data. 
 
```{r}
genre_peryear<-genre_year %>% 
  add_count(year,main_genre) %>% # making these a counts so ggplot can make them into bar plot
  add_count(year, name = "year_n") %>% 
  group_by(year,main_genre) # grouping for plotting
summary(genre_peryear) # another way to glimpse this data and give me min/maxs
```
 
# Time to plot

 I have never made a stacked bar plot that is a proportion. 

```{r genre_plot, dev= "png"}
ggplot(genre_peryear, aes(fill=main_genre,y=year_n, x=year))+
  geom_bar(position="fill", stat="identity") + # making this stacked as a %
  labs(title= "Genre Composition of True Horror films from 2000-2023", # adding a title
       x="Year", #changing title names 
       y="Proportion of Genre",
       fill="Genre")+ #changing legend names
  scale_y_continuous(expand = c(0,0),
                     labels= scales::percent)+ # axis will now be %
  scale_x_continuous(expand = c(0,0), 
                     limits =c(1999,2025), # had to choose 1999 so it would show 2000
                     breaks = seq(2000, 2024, by = 2))+ # all for readability
  scale_fill_manual(values=c("#34073d","#8179dd","#923e4d","#fb8654","#aec5a8"))+ #spooky season colors 
  theme_classic()+
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1)) # tilting values for readability 
```

 