---
title: "GoodPlotBadPlot"
author: "Megan Smith"
date: "2024-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

The figures and coding below will take you on a journey of how many hours a random selection of animals sleep.    
 **Caution, some figures below may cause anxiety and or discomfort**
 
  ---
#### Data

 The data being used is from the ggplot2 package called msleep and is an updated and expanded version of the mammals sleep dataset. Updated sleep times and weights were taken from V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007. 
 
  During this journey, we will soley looking at total sleep and awake times
  
##### First we will load in needed libraries 
```{r}
library(tidyverse)
library(here)
library(praise)
```

##### Next we will read in our data and glimpse it
```{r}
glimpse(msleep)
```

#### Next we will clean the data to remove NAs, select names less than 11 and also take a random selection of animals
```{r}
sleepNoNA<- msleep %>% 
  drop_na (sleep_total, name) #removing NAs in sleep_totals and name
sleep <- sleepNoNA %>%
  filter(nchar(name) < 11)
sleep <- sleep %>% 
  sample_n(size=35) # just selecting a random 35 animals 
```

Then we will make time awake and time asleep a percentage value
```{r}
sleep<- sleep %>% 
  mutate(per_sleep=(sleep_total/24)*100,
         per_awake = (awake/24)*100) %>% 
  pivot_longer(cols = c(per_sleep, per_awake),
               names_to = "day",
               values_to = "percentage") %>% 
  group_by(name) %>% 
  mutate(cumsum_per = cumsum(percentage),
         pos= percentage/ 2 + lag(cumsum_per, default = 0))
```


### CAUTION: THIS MAY CAUSE PAIN
 Making our bad plot
```{r}
ggplot(sleep, aes(x = "", y = percentage, fill = day)) +
  geom_bar(stat = "identity", width = 1.5) + # I heard bar plots are the best 
  coord_polar("y") +
  facet_wrap(~ name) +
  labs(title = "% of day each animal spends awake and asleep") +
  theme_void()+
  theme(plot.title = element_text(color = "yellow", size = 16, face = "bold"), # everyone loves yellow font
             legend.title = element_text(color = "lightblue"), #can you see this, I hope not
             legend.text = element_text(color = "lightblue"),
        plot.background = element_rect (fill="orange", color= NA),# wow this background hurts my eyes
        panel.background =  element_rect(fill= "tan", color = NA),# maybe this tan will help
        strip.text = element_text(size = 12, face= "bold")) + #again how can this be as unreadable 
  scale_fill_manual(values = c("per_awake" = "green", "per_sleep" = "red"), #worst colors ever
                    labels = c("per_awake" = "Awake", "per_sleep" = "Asleep")) + #changing label name
  geom_text(aes(y = pos, label = paste0(round(percentage, 1), "%")), color = "white") #making this as non-readable as possible

```


##### Congratulations! You made it through that 
```{r}
praise()
```


##### Now we will move onto a better representation of the data
```{r fig.width=8, fig.height=8}
ggplot(sleep, aes(x = reorder(name, -sleep_total), y = sleep_total)) +
  geom_segment(aes(xend = reorder(name, -sleep_total), yend = 0), #putting in a better order to read
               color = "#aeb6bf",
               size = 1, 
               alpha = 0.5) + 
  geom_point(color="#009688", size = 3, show.legend = TRUE) +  # point for hours slept
  labs(title="Total hours of sleep per day in animals",
       x="", y="Hours Slept",
       caption="Source: Msleep data from ggplot2") +
  theme_light() +
  theme(axis.title.y = element_blank(),
        axis.ticks.x = element_line(color = "#4a4e4d"),
        text = element_text(family = "sans", color = "#4a4e4d"), # who doesnt love sans
        strip.text.y.left = element_text(angle = 0),
        panel.background = element_rect(fill = "white", color = "white"),
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(color = "#4a4e4d", family = "sans"),
        plot.background = element_rect(fill = "white", color = "white"),
        panel.spacing = unit(0, "lines"),
        plot.margin = margin(1, 1, .5, 1, "cm")) +
   scale_y_continuous(expand = c(0, 0), limits= c(0,20)) +  # starting axis at 0 and making sure it goes until 20
   coord_flip()
```



